---
output:
  md_document:
    variant: markdown_github
---

```{r, include=FALSE, echo=FALSE}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
library(pacman)
p_load(fmxdat,tidyverse, rmsfuns, tbl2xts, PerformanceAnalytics, xts, glue, lubridate)

```
# Purpose

The aim of this folder is show my attempt to replicate two figures illustrating how different management fee structures effectively eat into investors overall cumulative returns. This involved first calculating the returns, starting with simple returns and then onto cumulative returns. Then calculating cumulative returns for different management fee structures, from 100bps to 350 bps. And finally 'pivoting longer' the returns for each different level of fees, in order to specify the colour option in the ggplot aesthetic, and finally plotting the cumulative returns. 

My attempt at this practical exercise, including all code and output used, can be found in the file 'Prac_1.md'.


# Cumulative Returns Since Inception
```{r, warning=FALSE, include=FALSE}

# Calculating Simple Returns

dta2 <- dta %>% arrange(date) %>%
    mutate(across(.cols = TRI, .fns = ~./lag(.) - 1, .names = "Simple_Returns"))%>%
# Specifying that the first date should be excluded (avoid NA's)
    filter(date>first(date))%>%
# Calculating cumulative returns
    mutate(Cum_Ret = cumprod(1 + Simple_Returns))

# calculating cumulative returns for different bps levels
# (100 bps approximately 1%)

dta22 <- dta2 %>%
  arrange(date) %>%
  mutate(bps_10=(Cum_Ret*0.999))%>%
  mutate(bps_50=(Cum_Ret*0.995))%>%
  mutate(bps_100=(Cum_Ret*0.99))%>%
  mutate(bps_200=(Cum_Ret*0.98))%>%
  mutate(bps_250=(Cum_Ret*0.975))%>%
  mutate(bps_350=(Cum_Ret*0.965))

# In order to create legend (ie by type), need to collect the bps columns into a single one (pivot longer)

dta221 <- dta22 %>%
  arrange(date)%>%
  pivot_longer(cols=starts_with("bps"), 
               names_to="Type",
               values_to="Tyranny" )
```

```{r, echo=FALSE, message=FALSE, warning =  FALSE, fig.align = 'left'}
    ggplot(dta221) + 
geom_line(aes(date, Tyranny, colour=Type)) + 
fmxdat::theme_fmx(title.size = ggpts(30), 
                    subtitle.size = ggpts(28),
                    caption.size = ggpts(25),
                    CustomCaption = T,
                  legend.size = 10,
                  legend.pos = "right") + 
    fmxdat::fmx_cols() + 
  
  labs(x = "", y = "Cumulative Returns", caption = "Note:\nOwn Calculations",
       title = "Cumulative Returns Since Inception",
       subtitle = "Base Return: FTSE JSE ALSI \nStart: June 2002")

```


# Cumulative Returns since 2010
```{r, warning=FALSE, include=FALSE}
dta3 <- dta221%>%filter(date>lubridate::ymd(20091231))

```

```{r, echo=FALSE, message=FALSE, warning =  FALSE, fig.align = 'left'}

    ggplot(dta3) + 
geom_line(aes(date, Tyranny, colour=Type)) + 
fmxdat::theme_fmx(title.size = ggpts(30), 
                    subtitle.size = ggpts(28),
                    caption.size = ggpts(25),
                    CustomCaption = T,
                  legend.size = 10,
                  legend.pos = "right") + 
    fmxdat::fmx_cols() + 
  
  labs(x = "", y = "Cumulative Returns", caption = "Note:\nOwn Calculations",
       title = "Cumulative Returns Since 2010",
       subtitle = "Base Return: FTSE JSE ALSI \nStart: January 2010")

```

# Attempting to correct for differences in cumulative returns
```{r, warning=FALSE, include=FALSE}

# Calculating Simple Returns

dta2 <- dta %>% arrange(date) %>%
    mutate(across(.cols = TRI, .fns = ~./lag(.) - 1, .names = "Simple_Returns")) %>%
# Specifying that the first date should be excluded (avoid NA's)
    filter(date>first(date)) %>%
  arrange(date) %>%
  mutate(bps_10=(Simple_Returns-(Simple_Returns*0.001))) %>%
  mutate(bps_50=(Simple_Returns-(Simple_Returns*0.005))) %>%
  mutate(bps_100=(Simple_Returns-(Simple_Returns*0.01))) %>%
  mutate(bps_200=(Simple_Returns-(Simple_Returns*0.02))) %>%
  mutate(bps_250=(Simple_Returns-(Simple_Returns*0.025))) %>%
  mutate(bps_350=(Simple_Returns-(Simple_Returns*0.035))) %>%
  
# ^ Calculating returns post-fees
  
# Calculating cumulative returns
        mutate(Cum_Ret1 = cumprod(1 + bps_10)) %>%
        mutate(Cum_Ret2 = cumprod(1 + bps_50)) %>%
        mutate(Cum_Ret3 = cumprod(1 + bps_100)) %>%
        mutate(Cum_Ret4 = cumprod(1 + bps_200)) %>%
        mutate(Cum_Ret5 = cumprod(1 + bps_250)) %>%
        mutate(Cum_Ret6 = cumprod(1 + bps_350)) %>%

pivot_longer(cols=starts_with("Cum"), 
               names_to="Type",
               values_to="Tyranny" )

head(dta2)

  
```

```{r, echo=FALSE, message=FALSE, warning =  FALSE, fig.align = 'left'}
    ggplot(dta2) + 
geom_line(aes(date, Tyranny, colour=Type)) + 
fmxdat::theme_fmx(title.size = ggpts(30), 
                    subtitle.size = ggpts(28),
                    caption.size = ggpts(25),
                    CustomCaption = T,
                  legend.size = 10,
                  legend.pos = "right") + 
    fmxdat::fmx_cols() + 
  
  labs(x = "", y = "Cumulative Returns", caption = "Note:\nOwn Calculations",
       title = "Cumulative Returns Since Inception",
       subtitle = "Base Return: FTSE JSE ALSI \nStart: June 2002")

```


I was, unfortunately, unable to replicate the figures exactly, the impact of higher management fees seems to be lower in the figures I plotted. Upon reflection, I realize that this may be due to the way I set up the 'Tyranny' column. Rather than subtracting the bps (for management fees), I should have calculated it as the percentage change in cumulative returns, after subtracting the management fees. ie the management fees should have been subtracted from the original return calculation, rather than the cumulative return calculation, and that is why the effect of greater management fees appears to be lower in the figures I was able to plot.

The second problem is more of a plotting issue, rather than a calculation issue. I was unable to replicate the floating caption in the figure specifying how much R1 million invested in year XYZ would've grown based on different management fee structures. 

This was a good opportunity to think more intuitively about how to approach this sort of plotting problem, and become more used to using the fmxdat auxiliary plotting functions.

    - Tiago

