---
output:
  md_document:
    variant: markdown_github
---

# Purpose

The aim of this folder is show my attempt to replicate two figures illustrating how different management fee structures effectively eat into investors overall cumulative returns. 

I was, unfortunately, unable to replicate the figures and calculate the impact on returns of different management fees. This tutorial however served as practice in using a project environment with the fmxdat package, using ggplot auxiliary functions, 


```{r}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
library(pacman)
p_load(tidyverse, rmsfuns, tbl2xts, PerformanceAnalytics, xts, glue, lubridate)

```


# Data loading

Loading in Jalshtr dataset from the fmxdat library.

```{r}

dta <- fmxdat::Jalshtr

# Checking missing values

sum(is.na(dta$TRI)) #=0... no missing TRI values
sum(is.na(dta$date)) #=0... no missing dates

```


# Cumulative Returns Since Inception
```{r, warning=FALSE}
# Calculating Simple Returns

dta2 <- dta %>% arrange(date) %>%
    mutate(across(.cols = TRI, .fns = ~./lag(.) - 1, .names = "Simple_Returns"))%>%
# Specifying that the first date should be excluded (avoid NA's)
    filter(date>first(date))%>%
# Calculating cumulative returns
    mutate(Cum_Ret = cumprod(1 + Simple_Returns))
    
# Plotting using fmxdat auxiliary functions
p<-
    ggplot(dta2) + 
geom_line(aes(date, Cum_Ret), color = "steelblue") + 
fmxdat::theme_fmx(title.size = ggpts(30), 
                    subtitle.size = ggpts(28),
                    caption.size = ggpts(25),
                    CustomCaption = T) + 
    fmxdat::fmx_cols() + 
  
  labs(x = "", y = "Cumulative Returns", caption = "Note:\nOwn Calculations",
       title = "Cumulative Returns Since Inception",
       subtitle = "Base Return: FTSE JSE ALSI \nStart: June 2002") + 
  
  guides(color = F)
  
  fmxdat::finplot(p, x.vert = T, x.date.type = "%Y", x.date.dist = "1 year")
```


# Cumulative Returns since 2010
```{r, warning=FALSE}
dta3 <- dta2%>%filter(date>lubridate::ymd(20100101))
p<-
    ggplot(dta3) + 
geom_line(aes(date, Cum_Ret), color = "steelblue") + 
fmxdat::theme_fmx(title.size = ggpts(30), 
                    subtitle.size = ggpts(28),
                    caption.size = ggpts(25),
                    CustomCaption = T) + 
    fmxdat::fmx_cols() + 
  
  labs(x = "", y = "Cumulative Returns", caption = "Note:\nOwn Calculations",
       title = "Cumulative Returns Since 2010",
       subtitle = "Base Return: FTSE JSE ALSI \nStart: January 2010") + 
  
  guides(color = F)
  
  fmxdat::finplot(p, x.vert = T, x.date.type = "%Y", x.date.dist = "1 year")

```

# Lastly

The point of this exercise was to investigate the "Tyranny of Fees"; how greater management fees impact investors cumulative returns. I have noted that a seperate column must be added for the "type" of management fee (ie the different scales) in order to change the color specification in the ggplot aesthetics. I look forward to learning how to be able to plot the figures, and continuing practicing using these tools in future tutorials. 
